Supporting Information : Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology

### 2019-03-15 MainText.Rmd
writing methods...continued

---
title: "Main text : Effect of copula on extinction risk"
author: "Shyamolina Ghosh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty

@@

@@

<!--density-dependence metapopulation model-->
We explored (Q2) by examining how nature of the tail dependence of an
environmental variable across space could influence the extinction risk of a 
metapopulation model with density-dependence feedback. We considered six density-dependent 
population-growth models (Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, Austin-Brewer)
in noisy environment (see Table \ref{tab_model_summary}) 
from a metapopulation context.
Models were adapted from earlier work
by @cohen1995unexpected, excluding only two : Varley model (this is not a continuous one) 
and Malthus-Condorcet-Mill model (terms containing logarithm of extinct population does not make sense). 
All the models could be represented via 
a generalized form : $\vec{P_{t+1}} = D \lambda_t \vec{P_t}$,
where $\vec{P_t}$ was a vector of population densities, 
the $i^{th}$ component $P_t^i$ representing
the density in the $i^{th}$ habitat patch ($i=1,\ldots,n$) at time $t$. 
The $n \times n$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
of stochastic density dependence form $\lambda_i(t)$ (see second column 
of Table \ref{tab_model_summary}). For simplicity, 
we considered $n$ patches in each model were arranged, evenly spaced, in a line
with the provision of "local" or "global" dispersal among patches. A n $\times$ n dispersal
matrix, $D$ was constructed assuming that a fraction $d$ of 
each population patch dispersed during any given time step, equally distributed to
the two or one nearest neighbors in case of local dispersal. For global dispersal, the fraction
$d$ that dispersed
from each patch was divided up equally among all other patches. We called $d$ as dispersal
rate with $d$ being zero there was no dispersal, and threby no chance to rescue an extinct 
patch through migration from nearby metapopulation patches.

The environmental noises from any two patch, say, 
$(\epsilon_t^i, \epsilon_t^j)$ were normally 
distributed within each patch with mean $m=0$ and standard deviation $sd$, 
(see the specified value $sd$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}), were 
independent through time, and showed the same 
spatial correlations between patches (to within sampling error) for every simulation. 
But they were made to exhibit right- or left-tail dependence between patches in
different simulations. A three dimensional array, numsims $\times$ n $\times$ numsteps, 
$\epsilon$, was made from these extreme tail-dependence noise 
matrix generated via algorithm stated in 
Appendix \ref{SM-extr_tail_dep} and used to calculate extinction risk 
for each of the specified model in Table \ref{tab_model_summary} for $numsteps = 25$ years, $n =$ 5 patches
and $numsims = 10000$ simulations.

<!--say how you set ext. threshold = Pe/10-->
To estimate extinction risk, we started with 
initial population $P_{t=0}^i = P_e^i$ for all patches ($i = 1,2,...,n$) where 
$P_e^i$ was the equilibrium value (or similar to carrying capacity concept) 
for each density-dependence model (see 4$^{th}$ column of Table \ref{tab_model_summary}).
We assumed population went extinct after each time step if it was below a certain threshold.
This extinction threshold was set to 10\% of initial population $P_e$ for individual model.

<!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->

<!--test hypothesis from Q3 : switching of ext risk with density-dependence metapopulation model with d=0-->

>>>>

---
title: "Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology"
author: "Shyamolina Ghosh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty

@@

\noindent\makebox[\linewidth]{\rule{\columnwidth}{1pt}}

\scriptsize

\textbf{ABSTRACT:}  \small write abstract here

\textbf{\textit{Keywords:}}

\noindent\makebox[\linewidth]{\rule{\columnwidth}{1pt}}

\normalsize

@@

<!--density-dependence metapopulation model-->
We explored (Q2) by examining how nature of the tail dependence of an
environmental variable across space could influence the extinction risk of a 
metapopulation model with density-dependence feedback. We considered six density-dependent 
population-growth models (Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, Austin-Brewer)
in noisy environment (see Table \ref{tab_model_summary}) 
from a metapopulation context.
Models were adapted from earlier work
by @cohen1995unexpected, excluding only two : Varley model (this is not a continuous one) 
and Malthus-Condorcet-Mill model (terms containing logarithm of extinct population does not make sense). 
All the models could be represented via 
a generalized form : $\vec{P}_{t+1} = D \lambda_t \vec{P_t}$,
where $\vec{P_t}$ was a vector of population densities, 
the $i^{th}$ component $P_t^i$ representing
the density in the $i^{th}$ habitat patch ($i=1,2,\ldots,n$) at time $t$. 
The $n \times n$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
of stochastic density dependence form $\lambda_i(t)$ (see second column 
of Table \ref{tab_model_summary}). For simplicity, 
we considered $n$ patches in each model were arranged, evenly spaced, in a line
with the provision of "local" or "global" dispersal among patches. A n $\times$ n dispersal
matrix, $D$ was constructed assuming that a fraction $d$ of 
each population patch dispersed during any given time step, equally distributed to
the two or one nearest neighbors in case of local dispersal. For global dispersal, the fraction
$d$ that dispersed
from each patch was divided up equally among all other patches. We called $d$ as dispersal
rate with $d$ being zero there was no dispersal, and threby no chance to rescue an extinct 
patch through migration from nearby metapopulation patches.

The environmental noises from any two patch, say, 
$(\epsilon_t^i, \epsilon_t^j)$ were normally 
distributed within each patch with mean $m=0$ and standard deviation $sd$, 
(see the specified value $sd$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}), were 
independent through time, and showed the same 
spatial correlations between patches (to within sampling error) for every simulation. 
But they were made to exhibit extreme right- or extreme left-tail dependence between patches in
different simulations. A three dimensional array, numsims $\times$ n $\times$ numsteps, 
$\epsilon$, was made from these extreme tail-dependence noise 
matrix generated via algorithm stated in 
Appendix \ref{SM-extr_tail_dep} and used to calculate extinction risk 
for each of the specified model in Table \ref{tab_model_summary} for $numsteps = 25$ years, $n =$ 5 patches
and $numsims = 10000$ simulations.

The models stated in Table \ref{tab_model_summary} had one thing in 
common that they showed undercompensatory
dynamics (monotonic approach towards stable steady state $P_e$) when
growth parameter $r$ remained below a critical 
level, say, $r_c$. If $r$ was beyond $r_c$, model showed 
overcompensatory dynamics with oscillatory approach 
towards stable equilibrium $P_e$. But if $r$ became too high and went
beyond $r_{bf}$, the parameter value responsible for Hopf-bifurcation,
model equilibrium $P_e$ became unstable and model could show sustained
oscillation and eventually led to chaos. Appendix \ref{SM-model_stability}
summarizes the stability analysis of each model in their deterministic form.

<!--say how you set ext. threshold = Pe/10-->
We started with $n$ patches of initial population $P_{t=0}^i = P_e^i, (i = 1,2,\ldots,n)$ where 
$P_e^i$ was the equilibrium value (or similar to carrying capacity concept) 
for each density-dependence model (see 4$^{th}$ column of Table \ref{tab_model_summary}).
We assumed population went extinct after each time step if it was below a certain threshold.
That extinction threshold was set to 10\% of initial population $P_e$ for individual model.

<!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
To test our hypothesis from (Q2) of \nameref{intro} for each metapopulation model,
we estimated extinction risk for a wide range of parameter $r$ 
which covered the undercompensatory and overcompensatory regime. We 
chose equispaced values of $r$ from the 95\% quantiles 
of possible range ($r_{min},r_{bf}$), where $r_{min}$ was the minimum value of $r$ 
for which the model equilibrium $P_e$ had always a finite, positive initial 
population (for details, see Appendix \ref{SM-model_stability}). For each chosen $r$ of each 
model where metapopulation could disperse locally with varying dispersal 
rate $d= 0,0.1,0.2,..0.5$, we calculated extinction risk with left 
tail dependent and right tail dependent environmental noise 
in different simulations, respectively. We also repeated the analysis 
for global dispersal for each metapopulation model. Script 
[`mydispersal.R`](https://github.com/sghosh89/ERC/blob/master/mydispersal.R) was used for this 
analysis.

<!--test hypothesis from Q3 : switching of ext risk with density-dependence metapopulation model with d=0-->
Given hypothesis from (Q2) was valid which scrutinized that whereas 
left tail dependence in noise made extinction risk 
higher in undercompensatory regime ($r < r_c$), it was the 
right tail dependent noise
which appeared to evoke higher extinction risk for metapopulations
in overcompensatory regime ($r > r_c$), we tested our third hypothesis for (Q3).
We considered
the switching to take place near $r_c$, not exactly at $r = r_c$ because the transition point $r_c$
from under to overcompensatory dynamics was expected to differ in stochastic environment than its 
deterministic analog depending on the underlying nonlinearity of the model.
We generated plots with extinction risk along y-axis
for both left tail and right tail dependent noises 
against the varying range of $r$ when no dispersal ($d=0$) was considered.
We excluded the possibility of dispersal in this case to avoid any rescue effect among patches. 
For each model with extreme left and extreme right tail dependent noises, we calculated 
2.5$^{th}$ and 97.5$^{th}$ quantiles
of 10000 simulations with the probability of computed
extinction risk at each $r$. We compared $r_c$ and the value of $r$ at which 
95\% CI of extinction risk from both noises overlapped
with each other. This could give us an idea for supporting our hypothesis.

### 2019-03-15 SuppMat.Rmd
writing methods...continued

---
title: "Supporting Information : Effect of copula on extinction risk"
author: "Shyamolina Ghosh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"

>>>>

---
title: "Supporting Information : Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology"
author: "Shyamolina Ghosh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"
